---
title: "Homework 1"
author: "Zichen Fan"
date: "2025-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

## Loading the data set 

```{r loading }
library(moderndive)

dat_weather=early_january_weather
```

## General information
```{r chunk_datainfo, eval = FALSE}
ncol(dat_weather)
nrow(dat_weather)

## Check important variables and value

skimr::skim(dat_weather)
```
The data set includes the weather and ambient data 
of EWR across 15 days in the first month of the year of 2013.

The data set includes 15 columns and 358 rows, with the following variables:

```{r chunk_variables, echo = FALSE}
names(dat_weather)

```
And the mean temperature of this data was 39.6 degree

## Temperature plot

```{r}
library(ggplot2)

temp_plot <- ggplot(data = dat_weather, 
              aes(x = time_hour, y = temp, color = humid ) )+
       geom_point()+
  
  ## name the plot and axis
       labs(title = " Temperature by hour ",
         x = "Time (Jan, 2013)",
         y = " Temperature (°F)",
         color = "Humidity") +
  
  ## adjust x axis labling 
          scale_x_datetime(
    date_breaks = "2 days",  
    date_labels = "%d")

temp_plot 
```

```{r, echo=FALSE}
ggsave(
  filename = "temp_vs_time_scatterplot.png", 
  plot = temp_plot)
```

Temperature shows daily fluctuations, generally rising during the day and falling at night, with an overall warming trend from January 2 to 14. Humidity tends to be higher during cooler periods and decreases as temperature increases over time, indicating an inverse relationship between humidity and temperature.

# Problem 2

## Create tibble

```{r, chunk_tibble}

library(tidyverse)

df_q2<-tibble(
  ## random sample of size 10 from a standard Normal distribution
    random_normal=rnorm(10),
    
  ## logical vector indicating whether elements of the sample are greater than 0
    positive=random_normal>0,
  
  ##character vector of length 10
    character_vec=letters[1:10],
  
  ##factor vector of length 10, with 3 different factor “levels”
    factor_vec=factor(rep(c("Low", "Medium", "High"), length.out = 10))
 )

df_q2
```

## Calculate mean

Try to calculate the mean of each variable. 
The result shows:

random_normal → ✅ can take mean (numeric)

positive → ✅ can take mean (logical; TRUE = 1, FALSE = 0)

character_vec → ❌ error (character: cannot compute mean)

factor_vec → ❌ error (factor: not numeric)
```{r, chunk_tibblemean}
mean(pull(df_q2, random_normal))

mean(pull(df_q2, positive))

mean(pull(df_q2,character_vec))

mean(pull(df_q2, factor_vec))
```


