---
title: "Homework 1"
author: "Zichen Fan"
date: "2025-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

## Loading the data set 

```{r loading }
library(moderndive)

dat_weather=early_january_weather
```

## General information
```{r chunk_datainfo, eval = FALSE}
ncol(dat_weather)
nrow(dat_weather)
skimr::skim(dat_weather)
```
The data set includes the weather and ambient data 
of EWR across 15 days in the first month of the year of 2013.

The data set includes 15 columns and 358 rows, with the following variables:

```{r chunk_variables, echo = FALSE}
names(dat_weather)

```
And the mean temperature of this data was 39.6 degree

## Temperature Plot

```{r}
library(ggplot2)

temp_plot <- ggplot(data = dat_weather, 
              aes(x = time_hour, y = temp, color = humid ) )+
       geom_point()+
       labs(title = " Temperature by hour ",
         x = "Time (Jan, 2013)",
         y = " Temperature (°F)",
         color = "Humidity") +
          scale_x_datetime(
    date_breaks = "2 days",  
    date_labels = "%d")
temp_plot 
```

```{r, echo=FALSE}
ggsave(
  filename = "temp_vs_time_scatterplot.png", 
  plot = temp_plot)
```

Temperature shows daily fluctuations, generally rising during the day and falling at night, with an overall warming trend from January 2 to 14. Humidity tends to be higher during cooler periods and decreases as temperature increases over time, indicating an inverse relationship between humidity and temperature.


